---
title: "Example_analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example_analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 0. Load libraries

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(LogConcDEAD)
```

## 0. Data Description

### Load Data from `TidyTuesday`

```{r message=FALSE, warning=FALSE}
if (!dir.exists("data")){
  dir.create("data")
  sports <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-03-29/sports.csv')
  write.csv(sports,"data/sports.csv")
}
sports <- read.csv("data/sports.csv")[,-1]
head(sports)
```

This data comes from [Equity in Athletics Data Analysis](https://ope.ed.gov/athletics/#/datafile/list), hattip to [Data is Plural](https://www.data-is-plural.com/archive/2020-10-21-edition/).

This data is consist of year, school, and many other variables related to collegiate sports in Unites States. Please see the Schoolsdoc file in each folder (EADA_2016-2017, etc) for the definitions of additional columns.

### Data Dictionary

# `sports.csv`

|variable             |class     |description |
|:--------------------|:---------|:-----------|
|year                 |integer   | year, which is year: year + 1, eg 2015 is 2015 to 2016 |
|unitid               |double    | School ID |
|institution_name     |character | School name |
|city_txt             |character | City name |
|state_cd             |character | State abbreviation |
|zip_text             |character | Zip of school |
|classification_code  |double    | Code for school classification |
|classification_name  |character | School classification |
|classification_other |character | School classification other |
|ef_male_count        |double    | Total male student  |
|ef_female_count      |double    | Total Female student  |
|ef_total_count       |double    | Total student  for binary male/female gender (sum of previous two cols) |
|sector_cd            |double    | Sector code |
|sector_name          |character | Sector name |
|sportscode           |double    | Sport code |
|partic_men           |double    | Participation men  |
|partic_women         |double    | Participation women |
|partic_coed_men      |double    | Participation as coed men |
|partic_coed_women    |double    | Participation for coed women|
|sum_partic_men       |double    | Sum of participation for men |
|sum_partic_women     |double    | Sum of participation women |
|rev_men              |double    | Revenue in USD for men |
|rev_women            |double    | Revenue in USD for women |
|total_rev_menwomen   |double    | Total revenue for both|
|exp_men              |double    | Expenditures in USD for men |
|exp_women            |double    | Expenditures in USD for women |
|total_exp_menwomen   |double    | Total Expenditure for both |
|sports               |character | Sport name |


## 1. Types of Sports

**Analysis Question** Q. What types of sports do students play the most in United States? and How are they distributed?

```{r}
sports_df <-sports %>%
  select(c("state_cd","total_rev_menwomen","total_exp_menwomen","sports")) %>%
  filter(!is.na(total_rev_menwomen),!is.na(total_exp_menwomen),!is.na(sports)) %>%
  rename(state = state_cd, revenue = total_rev_menwomen, expend = total_exp_menwomen) %>%
  mutate(sports = factor(sports))

sports_df %>%
  group_by(sports) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=count, y = reorder(sports, count))) +
  geom_bar(stat='identity') +
  xlab("Count") +
  ylab("Sports") +
  labs(title = "Types of Sports in US Schools", 
       subtitle = "from 2016 to 2017",
       caption = "In descending order of the counts of each sport type")
```

Basketball, Volleyball and Soccer are the top three sports reported in this data which show more than 8000 counts. 

## 2. Total Expenditures/Revenue for Each State

**Analysis Question** Q. Are there positive correlation between total expenditures and total revenue in sports for students in US schools?

```{r}
sports_df %>%
  split(.$state) %>%
  map(~ lm(revenue ~ expend, data=.)) %>%
  map(summary) %>%
  map_dbl("r.squared")
```

In terms of `r.squared` statistics in simple linear regression, We can observe that there is a significant linear correlation bewteen those two. 


```{r}
top10 <- sports_df %>%
  group_by(state) %>%
  summarise(count=n())%>%
  top_n(n=8, wt = count) %>%
  select("state")

sports_df %>% 
  filter(state %in% top10$state) %>%
  ggplot(aes(x=expend, y = revenue)) +
  geom_line()+
  facet_wrap(~state) +
  xlab("Total Expenditures in USD") +
  ylab("Total Revenue in USD") +
  labs(title = "Relationship between Expenditures and Revenue for sports in US", 
       subtitle = "by states with larger given dataset",
       caption = "There is a positive correlation between the expenditure and revenues")
```

For simpler visualization, we selected top 10 states with larger dataset and it definitely shows positive relationship between expenditures and revenue in sports for students in US schools.

## 3. MLE estimates for Expenditures/Revenue

**Analysis Question** Q. What is the estimated density of each expenditure and revenue for all collegiate sports in United States?

```{r}
sports_df %>% 
  filter(state== "MD", expend < 1000000) %>%
  ggplot(aes(x=expend)) +
  geom_histogram() +
  xlab("Expenditures") +
  ylab("Count") +
  labs(title = "Distribution of Expenditure(USD) in Collegial Sports", 
       subtitle = "in Maryland",
       caption = "It's mean is $196,205 and median is $114,522")
```

Distribution of Expenditures in Maryland Collegial sports is right-skewed with mean value $196,205. 

```{r warning=FALSE, message=FALSE}
MD <- sports_df %>% filter(state== "MD", expend < 200000, revenue < 200000)
hist(MD$expend)
hist(MD$revenue)
MD_df = cbind(MD$expend,MD$revenue)
lcd <- mlelcd(MD_df)
g <- interplcd(lcd)
plot(lcd, g=g, type="c", uselog=TRUE)
plot(lcd, g=g, type="i", uselog=TRUE)
```

We only selected data from Maryland data without too large values to avoid outliers and maintain log-concave distribution. 

We can observe how are 2-dimensional density of Expenditures/Revenue distributed using `LogConcDEAD`. It shows elliptical distribution centered in its mean value. 


## 4. List of Functions

### dplyr

-   filter()
-   mutate()
-   select()
-   rename()
-   group_by()
-   summarise()

### purrr

-   map()
-   map_dbl()

### ggplot2

-   geom_bar()
-   geom_line()
-   geom_histogram()
-   facet_wrap()

### LogConcDEAD

-   mlelcd()
-   interplcd()






